Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2015-09-12 3dpas_intro.tex
conclusion.

Many fields, such as biology, that until recently were not
characterized by their intensive use of data are rapidly becoming
data-driven.  The growth in data has profound implications for the
design and implementation of services and infrastructure.  For
example, there is a lifecycle to data that has often been ignored
until recently: the decreasing cost of storage has made it attractive
to archive \emph{everything}, just in case it becomes useful. This
approach is becoming untenable in many areas: some datasets are simply
too big to archive, and so must be summarized and their raw form
discarded (possibly in real time). Others are easier to regenerate
than to store.  Still others are too transient to justify their
storage, and may in aggregate be too big to store even when being
small individually.

Additionally, it is useful to observe that the trend toward more
unstructured and more dynamic data sources lead to higher demands on
data processing infrastructure.~\cite{snia_2012}
Any data science architecture needs to
explicitly recognize, capture, and manage the lifecycle requirements
of different datasets, within the constraints of available time and
storage. Associated with the growth in data volumes are increasing
levels of dynamism and distribution; sometimes distribution is
logical, but sometimes the data is physically distributed for reasons
of performance or convenience. At large scales, dynamism becomes
intrinsic to the application and infrastructure; sometimes it occurs
because of changing experimental or analytic conditions, sometimes due
to the availability of infrastructure and resources.
%We close out this paper with a brief
%discussion of a range of issues associated with the distribution and
%dynamism of data.

We have identified many different application approaches, programming
systems, and infrastructure techniques applicable to dynamic data.
Consistent with experience in other related fields, there seems no
reason to expect that these approaches and techniques will disappear
in the near future. While a convergence in approaches and techniques
might be expected, this has proved to be elusive; there remains a
diversity in the design, implementation, and execution of scientific
codes.  It may even be argued that it is premature to have convergence;
after all, D3 Science is still an emerging domain.

From the previous discussion we can draw out some conclusions about
the trends within dynamic, distributed, data-intensive science. For
instance, in \S\ref{sec:scenarios}, we divided applications into
groups: traditional and infrastructural applications.  Traditional
applications have been built by one or more authors in order to solve
a particular problem or to answer a science question.  Infrastructural
applications have been built by groups or communities to allow the
solution of a set of problems, or to answer a set of science
questions.  In both cases, we can observe that the same set of
applications stages (as shown in
\figurename~\ref{fig:figures_application-stages}) can be applied, with
APIs defined as the inputs and outputs of those stages.  These APIs
are very informal and loose for the traditional applications, but more
fixed and conservative for the infrastructural applications.
% \jhanote{is there a basis for these conjectures?} \katznote{just the
%   evidence of the applications themselves}
These APIs limit how new application paradigms %\jhanote{new paradigms of what?}
can be introduced for the infrastructural applications, unless those
paradigms happen to fit the existing APIs.  Another way of saying this
is that the abstractions for dynamism and distribution that can be
used in the infrastructural applications are those abstractions that
the developers of the original applications were aware of at the time
the infrastructural applications were defined.  This leads to a
possible mismatch, where computer scientists have new ideas that could
be applied to the class of problems that an infrastructural
application was designed to solve, but that the infrastructural
application doesn't allow, due to the fixed APIs.
New tools may be developed to address a
narrow set of features while ignoring the larger set of problems that
prevented earlier tools from being used.  These tools then may be used
be new applications as those applications are developed; however, they
are unlikely be used by the existing applications they were meant to
help.

This is because applications
(particularly infrastructural applications)
aren't developed in a vacuum, and that the constraints on the
applications lead to new tools that are or could be developed not
being able to be applied in practice.  For example, while
workflow systems are often sufficiently domain-independent that
there does not seem to be a tie of a particular system to a particular
infrastructural application, there are a very large number of workflow
systems that exist, and it is common in the authors' experience to
hear about new ones being developed from scratch because ``the
ones that exist don't quite do what is needed for my application or system.''

@@

As part of this survey, we also observe that there exist patterns in
the described applications; however, implementation and deployment
specific details can often mask these patterns.  For example,
applications often differ in the way they handle distribution and
dynamism, such as managing such as data/compute localities. This makes
it difficult to support patterns of distribution and dynamism in a
general-purpose fashion, for anything but the simplest patterns. In
\S\ref{sec:quantifyingdistapps} we have attempted to initiate a
discussion in this context however by considering two
ratios %(identified as parameters $\alpha$ and $\beta$)
to relate data volume and distribution with compute operations
performed by an application. We believe determining values for these
parameters would provide a useful starting point for comparing
infrastructure requirements of Big Data applications.
%\jhanote{I transposed the last two paragraphs}

@@

\section*{Acknowledgements}
This work was sponsored by NSF OCI-1059635 as well as made possible by the UK
e-Science Institute sponsored Research Theme on Dynamic Distributed
Data-Intensive Programming Abstractions and Systems. We thank Jon Weissman and
Jon Blower for useful initial discussions, as well as 3DPAS workshop
attendees. Discussions of the selected applications benefited from the generous
help of Adam Barker, Jon Blower, Julian Borrill, Silvia Delgado Olabarriaga,
Steve Fisher, Keith Jackson, Scott Klasky, Bob Mann, Don Middleton, Kees
Nieuwenhuis, Manish Parashar, Stephen Pascoe, and Jon Weissman, though any
errors are the responsibility of the authors.  Some of the work by Katz was
supported by the National Science Foundation while working at the Foundation;
any opinion, finding, and conclusions or recommendations expressed in this
material are those of the author(s) and do not necessarily reflect the views of
the National Science Foundation.

>>>>

The growth in data volumes and its importance is having profound implications on
the way applications are designed to the way we develop infrastructure and
provision associated services.  Many fields, such as biology, that until
recently were not characterized by their intensive use of data are rapidly
becoming data-driven.

Associated with the growth in data volumes are increasing levels of dynamism and
distribution; sometimes distribution is logical, but sometimes the data is
physically distributed for reasons of performance or convenience. At large
scales, dynamism becomes intrinsic to the application and infrastructure;
sometimes it occurs because of changing experimental or analytic conditions,
sometimes due to the availability of infrastructure and resources. In this
paper, we have identified many different applications characteristics,
programming systems, and infrastructure techniques that either support dynamic
data or produce it.

% For example, there is a lifecycle to data that has often been ignored until
% recently: the decreasing cost of storage has made it attractive to archive
% \emph{everything}, just in case it becomes useful. This approach is becoming
% untenable in many areas: some datasets are simply too big to archive, and so
% must be summarized and their raw form discarded (possibly in real time). Others
% are easier to regenerate than to store.  Still others are too transient to
% justify their storage, and may in aggregate be too big to store even when being
% small individually.  Additionally, it is useful to observe that the trend toward
% more unstructured and more dynamic data sources lead to higher demands on data
% processing infrastructure~\cite{snia_2012}. Any data science architecture needs
% to explicitly recognize, capture, and manage the lifecycle requirements of
% different datasets, within the constraints of available time and storage.

% Consistent with experience in other related fields, there seems no reason to
% expect that these approaches and techniques will disappear in the near
% future. While a convergence in approaches and techniques might be expected, this
% has proved to be elusive; there remains a diversity in the design,
% implementation, and execution of scientific codes.  It may even be argued that
% it is premature to have convergence; after all, D3 Science is still an emerging
% domain.

% From the previous discussion we can draw out some conclusions about the trends
% within dynamic, distributed, data-intensive science. For instance,

In \S\ref{sec:scenarios}, we divided applications into groups: traditional and
infrastructural applications.  Traditional applications have been built by one
or more authors in order to solve a particular problem or to answer a science
question.  Infrastructural applications have been built by groups or communities
to allow the solution of a set of problems, or to answer a set of science
questions. All applications in general, but infrastructural applications in
particular cannot be developed in vacuum; they must be executed on shared
infrastructure, and are thus sensitive to the external tools and services, as
well as their provisioning as software-systems.

As part of this survey, we observe that there exist patterns in the described
applications; however, implementation and deployment specific details can often
mask these patterns.  For example, applications often differ in the way they
handle distribution and dynamism, such as managing such as data/compute
localities. This makes it difficult to support patterns of distribution and
dynamism in a general-purpose fashion, for anything but the simplest patterns.

% In both cases, we can observe that the same set of applications stages (as shown
% in \figurename~\ref{fig:figures_application-stages}) can be applied, with APIs
% defined as the inputs and outputs of those stages.  These APIs are very informal
% and loose for the traditional applications, but more fixed and conservative for
% the infrastructural applications. These APIs limit how new application
% paradigms can be introduced for the infrastructural applications, unless those
% paradigms happen to fit the existing APIs.  Another way of saying this is that
% the abstractions for dynamism and distribution that can be used in the
% infrastructural applications are those abstractions that the developers of the
% original applications were aware of at the time the infrastructural applications
% were defined.  This leads to a possible mismatch, where computer scientists have
% new ideas that could be applied to the class of problems that an infrastructural
% application was designed to solve, but that the infrastructural application
% doesn't allow, due to the fixed APIs.  New tools may be developed to address a
% narrow set of features while ignoring the larger set of problems that prevented
% earlier tools from being used.  These tools then may be used by new applications
% as those applications are developed; however, they are unlikely be used by the
% existing applications they were meant to help.

% constraints on the applications lead to new tools that are or could be developed
% not being able to be applied in practice.

% For example, while workflow systems are often sufficiently domain-independent
% that there does not seem to be a tie of a particular system to a particular
% infrastructural application, there are a very large number of workflow systems
% that exist, and it is common in the authors' experience to hear about new ones
% being developed from scratch because ``the ones that exist don't quite do what
% is needed for my application or system.''

@@

% In \S\ref{sec:quantifyingdistapps} we have attempted to initiate a discussion in
% this context however by considering two ratios to relate data volume and
% distribution with compute operations performed by an application. We believe
% determining values for these parameters would provide a useful starting point
% for comparing infrastructure requirements of data-intensive applications and
% systems.

%\jhanote{I transposed the last two paragraphs}

@@

\section*{Acknowledgements}
This work was sponsored by NSF OCI-1059635 and made possible by the UK e-Science
Institute Research Theme on Dynamic Distributed Data-Intensive Programming
Abstractions and Systems. We thank Jon Weissman and Jon Blower for useful
initial discussions, as well as 3DPAS workshop attendees. Discussions of the
selected applications benefited from the generous help of Adam Barker, Jon
Blower, Julian Borrill, Silvia Delgado Olabarriaga, Steve Fisher, Keith Jackson,
Scott Klasky, Bob Mann, Don Middleton, Kees Nieuwenhuis, Manish Parashar,
Stephen Pascoe, and Jon Weissman, though any errors are the responsibility of
the authors.  Some of the work by Katz was supported by the National Science
Foundation while working at the Foundation; any opinion, finding, and
conclusions or recommendations expressed in this material are those of the
author(s) and do not necessarily reflect the views of the National Science
Foundation.
